Staged Abstract Interpreters

### 2019-09-11 paper/80_related.tex
misc
typo

Splitting an analysis into multiple stages has also been studied for other
forms of static analyses.
\citet{DBLP:conf/cgo/HardekopfL11} applied staging to a
flow-sensitive pointer analysis. The first stage is to analyze the
programs and obtain a sparse representation, and then the second
stage conducts a flow-sensitive analysis by refining the first
one. Many other forms of static analysis are implemented by
first collecting constraints from programs and then using a solver to discharge
the proof obligations or find the fixed-point. Notable examples include SAT-
and SMT-based program analysis
\cite{Gulwani:2008:PAC:1375581.1375616} and Datalog-based points-to
analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
Such analyses can be considered having stage distinctions.  One piece of future
work is to explore whether they can be implemented and optimized using
meta-programming.
Souffl{\'e}, one of the state-of-the-art Datalog engines commonly used in
points-to analysis \cite{Antoniadis:2017:PDS:3088515.3088522}, uses Futamura
projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
However, it only specializes the solving engine to Datalog rules, instead of a
full pipeline from a program AST to a specialized solver.

@@

\paragraph{Two-level Semantics} The idea of reinterpreting the semantics as
abstract interpretation can be traced to Nielson's two-level semantics
\cite{NIELSON1989117}; using two-level semantics for code generation
was also explored by \citet{NIELSON198859}.
\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract
interpreters is closely related to the two-level semantics: the
use of a generic interface with monads and then reinterpreting it with
different semantics is already two-level. Instead of focusing on
semantics, this paper shows how a staged analyzer can be built and
used to increase efficiency of static analyses. We augment the monadic
abstract interpreter with binding-time polymorphism, using code generation to
produce efficient low-level code. The presented work can be considered as
a two-dimensional application of the two-level semantics.

>>>>

Splitting an analysis into multiple stages has also been studied for other
forms of static analysis.
\citet{DBLP:conf/cgo/HardekopfL11} applied staging to a
flow-sensitive pointer analysis. The first stage is to analyze the
programs and obtain a sparse representation, and then the second
stage conducts a flow-sensitive analysis by refining the first
one. Many other forms of static analysis are implemented by
first collecting constraints from programs and then using a solver to discharge
the proof obligations or find the fixed-point. Notable examples include SAT-
and SMT-based program analysis
\cite{Gulwani:2008:PAC:1375581.1375616} and Datalog-based points-to
analysis \cite{Smaragdakis:2015:PA:2802194.2802195}.
Such analyses can be considered having stage distinctions.  One piece of future
work is to explore whether they can be implemented and optimized using
meta-programming.
Souffl{\'e}, one of the state-of-the-art Datalog engines commonly used in
points-to analysis \cite{Antoniadis:2017:PDS:3088515.3088522}, uses Futamura
projections for efficient implementation \cite{10.1007/978-3-319-41540-6_23}.
However, it only specializes the solving engine to Datalog rules, instead of a
full pipeline from a program AST to a specialized solver.

@@

\paragraph{Two-level Semantics} The idea of reinterpreting the semantics as
abstract interpretation can be traced to Nielson's two-level semantics
\cite{NIELSON1989117}; using two-level semantics for code generation
was also explored by \citet{NIELSON198859}.
\citet{Sergey:2013:MAI:2491956.2491979}'s monadic abstract
interpreter is closely related to the two-level semantics: the
use of a generic interface with monads and then reinterpreting it with
different semantics is already two-level. Instead of focusing on
semantics, this paper shows how a staged analyzer can be built and
used to increase efficiency of static analyses. We augment the monadic
abstract interpreter with binding-time polymorphism, using code generation to
produce efficient low-level code. The presented work can be considered as
a two-dimensional application of the two-level semantics.
